Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-02-03 manuscript/manuscript.Rmd
Update manuscript.Rmd

- Set weights based on quantiles of the original feature $x$. Quantiles are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.

@@

- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (x)$ and the theoretical z-score from a standard normal distribution: $\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.

Based on either of the above, weights are set using a weight function. We use the following functions:

- A step function, with $k_1$ as threshold parameter:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| x_i \right| \leq k_1\\
0 & \text{if } \left| x_i \right| > k_1
\end{cases}
\end{equation}
```

- A triangle function (or generalised Huber weight), with $k_1$ and $k_2 \geq k_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| x_i \right| \leq k_1\\
 \\
0 & \text{if } \left| x_i \right| > k_1
\end{cases}
\end{equation}
```

@@

@@

```{r, include=FALSE, eval = FALSE}
# Generator for creating parameters and data for all experiments.
generate_experiment_data <- coro::generator(
  function(
    x_normal,
    x_right_skewed,
    x_left_skewed){

>>>>

- Set weights based on probabilities of the empirical distribution of the original feature $x$. Probabilities are determined as $p_i = \frac{i - 1/3}{n + 1/3}$, with $i = 1, 2, \ldots n$, with $n$ the number of instances of feature $x$.

@@

- Set weights based on the residual error between the z-score of the transformed feature $\phi^{\lambda, x_0} (x)$ and the theoretical z-score from a standard normal distribution: $r_i =\left| \left( \phi^{\lambda, x_0}(x_i) - \mu_M)\right) / \sigma_M - F^{-1}_{\mathcal{N}}(p_i) \right|$, with $\mu_M$, $\sigma_M$ and $p_i$ as defined above.

Based on either of the above, weights are set using a weighting function. We will investigate the following functions, where $x_i$ can be either $p^{*}_i=2 \left( p_i - 0.5\right)$ (so that the quantiles are zero-centred), $z_i$ or $r_i$:

- A step function, with $k_1 \geq 0$ as threshold parameter:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| x_i \right| \leq k_1\\
0 & \text{if } \left| x_i \right| > k_1
\end{cases}
\end{equation}
```

- A triangle function (or generalised Huber weight), with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| x_i \right| < k_1\\
1 - \frac{\left| x_i \right| - k_1}{k_2 - k_1} & \text{if } k_1 \leq \left| x_i \right| \leq k_2 \\
0 & \text{if } \left| x_i \right| > k_2
\end{cases}
\end{equation}
```

- A tapered cosine function [@Tukey1967-eb], with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| x_i \right| < k_1\\
0.5 + 0.5 \cos\left(\pi \frac{\left| x_i \right| - k_1}{k_2 - k_1} \right) & \text{if } k_1 \leq \left| x_i \right| \leq k_2 \\
0 & \text{if } \left| x_i \right| > k_2
\end{cases}
\end{equation}
```

All weighting functions share the characteristic that for $x_i < k_1$, instances are fully weighted, i.e. when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top. The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq x_i \leq k_2$, and assign no weight to instances $x_i > k_2$.

```{r weight-functions , echo=FALSE, warning=FALSE, message=FALSE, fig.cap=cap, eval = FALSE}
cap <- paste0(
  "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
  "The step function was parameterised with $k_1 = 0.60$. ",
  "The triangle and tapered cosine functions were both parameterised with $k_1 = 0.30$ and $k_2 = 0.90$.")

data_step <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
data_step[, "weight":=power.transform:::.step_window(x, k1=0.60)]

p_step <- ggplot2::ggplot(
  data = data_step,
  mapping = ggplot2::aes(
    x = x,
    y = weight))
p_step <- p_step + plot_theme
p_step <- p_step + ggplot2::geom_line()
p_step <- p_step + ggplot2::ggtitle(label="step")
p_step <- p_step + ggplot2::theme(
  plot.title=ggplot2::element_text(
    hjust=0.5,
    size=ggplot2::rel(1.0)))

data_triangle <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
data_triangle[, "weight":=power.transform:::.triangular_window(x, k1=0.30, k2=0.90)]

p_triangle <- ggplot2::ggplot(
  data = data_triangle,
  mapping = ggplot2::aes(
    x = x,
    y = weight))
p_triangle <- p_triangle + plot_theme
p_triangle <- p_triangle + ggplot2::geom_line()
p_triangle <- p_triangle + ggplot2::ggtitle(label="triangle")
p_triangle <- p_triangle + ggplot2::theme(
  plot.title=ggplot2::element_text(
    hjust=0.5,
    size=ggplot2::rel(1.0)),
  axis.text.y = ggplot2::element_blank(),
  axis.title.y = ggplot2::element_blank())

data_cosine <- data.table::data.table(x = seq(from=-1.0, to=1.0, by=0.001))
data_cosine[, "weight":=power.transform:::.tapered_cosine_window(x, k1=0.30, k2=0.90)]

p_cosine <- ggplot2::ggplot(
  data = data_cosine,
  mapping = ggplot2::aes(
    x = x,
    y = weight))
p_cosine <- p_cosine + plot_theme
p_cosine <- p_cosine + ggplot2::geom_line()
p_cosine <- p_cosine + ggplot2::ggtitle(label="tapered cosine")
p_cosine <- p_cosine + ggplot2::theme(
  plot.title = ggplot2::element_text(
    hjust = 0.5,
    size = ggplot2::rel(1.0)),
  axis.text.y = ggplot2::element_blank(),
  axis.title.y = ggplot2::element_blank())

p <- egg::ggarrange(plots=list(p_step, p_triangle, p_cosine), ncol=3, draw=FALSE)
p
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, eval = FALSE}

set.seed(95)

.compute_lambda <- function(x){
  # Create transformer object.
  transformer <- suppressWarnings(
    power.transform::find_transformation_parameters(
      x = x,
      method = "yeo_johnson",
      robust = FALSE,
      shift = TRUE))

  return(transformer@lambda)
}

# Generate table of asymmetric generalised normal distribution parameters.
n_distributions <- 100

# Generate alpha, beta and n.
n <- stats::runif(n=n_distributions, min=2, max=4)
n <- round(10^n)

alpha <- stats::runif(n=n_distributions, min=0.01, max=0.99)
beta <- stats::runif(n=n_distributions, min=1.00, max=5.00)

# Generate corresponding distributions.
x <- mapply(
  power.transform::ragn,
  n = n,
  alpha = alpha,
  beta = beta)

# Compute lambda values without weighting for Yeo-Johnson.
target_lambda <- sapply(
  x,
  .compute_lambda)

# Add outliers, between 0.00 and 0.10.
n_outliers <- 10
outlier_fraction <- (seq_len(n_outliers) - 1)^2 /((n_outliers-1)^2 * 10)

# Find optimised k1 and k2 values for all combinations of weight sources and
# weighting functions.
x <- mapply(
  function(x, n, alpha, beta, target_lambda, outlier_fraction){

    parameters <- list("n" = n, "alpha"=alpha, "beta" = beta, "target_lambda"=target_lambda)

    x <- lapply(
      outlier_fraction,
      function(k, x, parameters){
        # Set parameters.
        parameters$k <- k

        # Compute interquartile range.
        interquartile_range <- stats::IQR(x)

        return(list(
          "x" = x_outlier,
          "parameters" = parameters))
      },
      x = x,
      parameters)

    return(x)
  },
  x = x,
  n = n,
  alpha = alpha,
  beta = beta,
  target_lambda = target_lambda,
  MoreArgs=list("outlier_fraction"=outlier_fraction),
  SIMPLIFY=FALSE,
  USE.NAMES=FALSE)

x <- unlist(x, recursive = FALSE)

# Optimise k1, k2 so that sum of absolute differences with target-lambda is
# minimised for each combination of source and weighting method.

@@

```{r, eval = FALSE}
# Generator for creating parameters and data for all experiments.
generate_experiment_data <- coro::generator(
  function(
    x_normal,
    x_right_skewed,
    x_left_skewed){
